---
title: "Assignment 4 - Task 2"
author: "Samantha Kuglen"
date: "11/5/2021"
output: html_document
---
## 3A) Overview

**Data Citation: **

## 3B) Data and Analysis

## Setup
```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(ggplot2)
```

## Read in the Data
```{r}
# Store Bonanza Hares csv as new object `hare_data`, update date to YYYY-MM-DD format, and create new column `year`
hare_data <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  mutate(site_names = case_when(
    grid %in% c("bonrip") ~ "Bonanza Riparian",
    grid %in% c("bonmat") ~ "Bonanza Mature",
    grid %in% c("bonbs") ~ "Bonanza Lowland Black Spruce"
  ))
```

## 3C - I: Juvenile Hare Initial Investigation
```{r}
# Count the total # of juvenile hare trappings during each year of the study

juvenile_hare_counts <- hare_data %>%  
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(age =="j") %>% 
  group_by(year) %>% 
  summarize(count = n()) %>% 

juvenile_hare_stats <- hare_data %>% 
  filter(age =="j") %>% 
  summarize(hare_count = n(),
            mean = mean(hare_count, na.RM = TRUE,),
            median = median(hare_count, na.RM = TRUE),
            sd = sd(hare_count),
            max = max(hare_count),
            min = min(hare_count))
  
```

## 3C - I: Juvenile Hare Data Visualization
```{r, fig.cap='cap'}
# Finalized data visualization of the juvenile hare counts by year
ggplot(yearly_hare_trappings, aes(x = year)) +
  geom_histogram() + 
  scale_x_continuous(1998:2012)
```
Major takeaways from the graph about total snowshoe hare trappings: 
- The minimum and maximum counts of juvenile hare trappings
- Mean and median annual number of juvenile hares trapped
- Notable trends include
- Moving forward, to standardize the juvenile hare population in future analyses

## 3C - II: Visualize Juvenile Hare Weights
```{r, fig.cap='cap'}
# Finalized data visualizations comparing juvenile hare weights by sex and site (grid) and omitting other variables and including groups where variables are not recorded (`NA`)
```

Main takeaways from the graph: 

## 3C - III: Juvenile Weight Comparison (Male & Female Snowshoe Hares)
Further comparison of mean weights for **juvenile male and female snowshoe hares**
```{r}
# Finalized table containing the mean, standard deviation, and sample size for male and female juvenile snowshoe hares, with a table caption
```
Comparison of the means: 
- The actual difference in means
- Effect size (Cohen's *d*)
- Outcome of an unpaired two-sample t-test

## 3C - IV: Relationship Between Juvenile Weight & Hind Foot Length
Exploration of the relationship between juvenile snowshoe hare hind foot length (mm) and weight of all juvenile hares in the study.
```{r}
# Final visualization of juvenile hare hind foot length versus weight.
```


