---
title: "Assignment 4 - Task 2"
author: "Samantha Kuglen"
date: "11/5/2021"
output: html_document
---
### 3A. Overview
This report is an exploratory review of counts and sizes of juvenile snowshoe hares (*Lepus americanus*), recorded at the [Bonanza Creek Experimental Forest (US Long Term Ecological Research Site)](https://www.lter.uaf.edu/) from 1998 - 2012. This report provides analyses on juvenile hare 1) annual trap counts, 2) weight comparison and 3) relationship between weight and hind foot length. This data is important because snowshoe hares are a keystone species in northern boreal forests. The purpose of this analysis is to explore reasons that this species may experience population fluctuations.

### 3B. Data and Analysis
- Describe Data

#### Statistical Methods and Tools

**Data Citation: ** Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14


```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(ggplot2)
library(effsize)
```

```{r read in data}
# Store Bonanza Hares csv as new object `hare_data`, update date to YYYY-MM-DD format, create new column `year`, add corresponding sex and site names 
juvenile_hares <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  filter(age == "j") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_names = case_when( # add new column for site names
    grid %in% c("bonrip") ~ "Bonanza Riparian",
    grid %in% c("bonmat") ~ "Bonanza Mature",
    grid %in% c("bonbs") ~ "Bonanza Black Spruce" 
  )) %>% 
  mutate(sex = case_when( # update coluimn for sex
    sex %in% c("m") ~ "Male",
    sex %in% c("f") ~ "Female",
    sex %in% c("NA") ~ "NA"
  ))
```


```{r}
# Count the total # of juvenile hare trappings during each year of the study
hare_counts <- juvenile_hares %>% 
  group_by(year) %>% 
  summarize(sample_size = n()) 

# Mean, median, max, min, of juvenile hares trapped
hare_stats <- hare_counts %>% 
  summarize(mean = mean(sample_size),
            median = median(sample_size),
            sd = sd(sample_size),
            max = max(sample_size),
            min = min(sample_size)) 
  
```

### 3C. Preliminary results
#### I: Data visualization of annual juvenile hare trap counts 
```{r, fig.cap='cap'}
# Finalized data visualization of the juvenile hare counts by year
ggplot(hare_counts, aes(x = year, y = sample_size)) +
  geom_col(color = "#037bfc", 
           fill = "#037bfc",
           alpha = 0.75) +
  scale_x_continuous(breaks = 1999:2012) + 
  scale_y_continuous(breaks = seq(min(0), max(130), by = 10)) +
  labs(x = "Year",
       y = "Number of Hares",
       title = "Annual Juvenile Hare Trap Counts (KUGLEN)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")

```
Major takeaways from the graph about total snowshoe hare trappings: 
- The minimum and maximum counts of juvenile hare trappings
- Mean and median annual number of juvenile hares trapped
- Notable trends include
- Moving forward, to standardize the juvenile hare population in future analyses

#### II: Comparison of juvenile hare weights
Finalized data visualizations comparing juvenile hare weights by sex and site (grid) and omitting other variables and including groups where variables are not recorded (`NA`)
```{r, fig.cap='cap'}
ggplot(juvenile_hares) + 
  geom_histogram(alpha = 0.6, aes(x = weight, fill = sex, color = sex)) + 
  facet_wrap(~site_names) +
  scale_fill_manual(values = c("#de5b5b", "#5bb9de", "#6c7375")) +
  scale_color_manual(values = c("#de5b5b", "#5bb9de", "#6c7375")) +
  labs(x = "Weight (g)",
       y = "Count",
       title = "Juvenile Snowshoe Hare Weight Comparison (KUGLEN)",
       subtitle = "1999 - 2012") +
  theme(plot.title = element_text(hjust = 0.5))
```

Main takeaways from the graph: 

#### III: Descriptive statistics of juvenile weight comparison 
Finalized table containing the mean, standard deviation, and sample size for male and female juvenile snowshoe hares, with a table caption
```{r}
weight_stats <- juvenile_hares %>% 
    group_by(sex) %>% 
    select(sex, weight) %>% 
    summarize(mean = mean(weight, na.rm = TRUE),
              sd = sd(weight, na.rm = TRUE),
              median = median(weight, na.rm = TRUE),
              hare_sample_size = n())

```
Comparison of the means: 
- The actual difference in means
- Effect size (Cohen's *d*)
- Outcome of an unpaired two-sample t-test

#### IV: Relationship between juvenile weight & hind foot length
Exploration of the relationship between juvenile snowshoe hare hind foot length (mm) and weight of all juvenile hares in the study.
```{r}
# Final visualization of juvenile hare hind foot length versus weight.
ggplot(juvenile_hares, aes(x = hindft, y = weight)) + 
  geom_point() + 
  labs(x = "Hind Foot Length (mm)",
       y = "Weight",
       title = "Juvenile Weight Comparison")

```

Linear Regression to describe how juvenile snowshoe hare weight changes with respect to hind foot length. 
```{r linear regression }
lm_hares <- lm(weight ~ hindft, data = juvenile_hares)
plot(lm_hares)

male_hares <- juvenile_hares %>% 
  filter(sex == "Male", age == "j") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)

female_hares <- juvenile_hares %>% 
  filter(sex == "Female", age == "j") %>% 
  filter(!is.na(weight)) %>% 
  pull(weight)

cohen_hares <- cohen.d(female_hares, male_hares)

t.test(male_hares, female_hares)
 
#female_mean <- mean(female_hares)
#male_mean <- mean(male_hares) 

#actual_mean_difference <- (male_mean - female_mean)

```

### Summary
